Trajectories

### 2021-09-15 Rmd/report.Rmd
Add production PPCs

# import data
tar_load(items)
tar_load(participants)
tar_load(vocabulary)
tar_load(responses)
tar_load(fits)

@@

  plot_layout(guides = "collect") &
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

## Prior-predictive checks

```{r ppc, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

@@

m <- add_fitted_draws(nd, fits$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

```{r fix_effets_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

post <- gather_draws(fits$fit_2, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

ggplot(post, aes(.value)) +
  facet_wrap(~.variable, scales = "free_x") +
  geom_vline(xintercept = 0, size = 1, colour = "grey") +
  stat_histinterval(fill = "lightblue") +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_github() +
  theme(
    legend.position = "none",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

### Participant

```{r random_participant, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  left_join(distinct(fits$fit_2$data, id, lp))

@@

### Item

```{r re_item, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fits$fit_2$data, te, lp))

@@

```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

m <- add_fitted_draws(nd, fits$fit_2, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

@@

  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "right",
    legend.title = element_blank()
  )

@@

```{r diagnostics_traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}

gather_draws(fits$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )

>>>>

# import data
tar_load(items)
tar_load(participants)
tar_load(vocabulary)
tar_load(responses)
tar_load(fits_comp)
tar_load(fits_prod)

@@

  plot_layout(guides = "collect") &
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point(size = 3) +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

## Prior-predictive checks (PPC)

### PPC: Comprehension

```{r ppc_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

@@

m <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

### PPC: Production

```{r ppc_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

m <- add_fitted_draws(nd, fits_prod$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Production)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

### Fixed effects: Compehension

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits_comp$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

```{r fix_effets_comp_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

post <- gather_draws(fits_comp$fit_2, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

ggplot(post, aes(.value)) +
  facet_wrap(~.variable, scales = "free_x") +
  geom_vline(xintercept = 0, size = 1, colour = "grey") +
  stat_histinterval(fill = "white") +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_github() +
  theme(
    legend.position = "none",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

### Fixed effects: Production

```{r fix_effets_prod_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

post <- gather_draws(fits_prod$fit_2, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

ggplot(post, aes(.value)) +
  facet_wrap(~.variable, scales = "free_x") +
  geom_vline(xintercept = 0, size = 1, colour = "grey") +
  stat_histinterval(fill = "white") +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_github() +
  theme(
    legend.position = "none",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

@@

### Random effects: Comprehension

#### Participant

```{r random_participant_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits_comp$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
  left_join(distinct(fits_comp$fit_2$data, id, lp))

@@

#### Item

```{r re_item_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits_comp$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_comp$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fits_comp$fit_2$data, te, lp))

### Random effects: Production

#### Participant

```{r random_participant_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits_prod$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
  left_join(distinct(fits_prod$fit_2$data, id, lp))

#### Item

```{r re_item_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits_prod$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits_prod$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fits_prod$fit_2$data, te, lp))

@@

### Marginal means: Comprehension

```{r emmeans_comp, message=FALSE, warning=FALSE, paged.print=FALSE}

@@

m <- add_fitted_draws(nd, fits_comp$fit_2, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

@@

  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

## Marginal means: Production

```{r emmeans_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

m <- add_fitted_draws(nd, fits_prod$fit_2, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

```{r diagnostics_traceplots_comp, message=FALSE, warning=FALSE, paged.print=FALSE}
gather_draws(fits_comp$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Comprehension") +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )
```

```{r diagnostics_traceplots_prod, message=FALSE, warning=FALSE, paged.print=FALSE}

gather_draws(fits_prod$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain", title = "Production") +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )
